Constructive forcing

### 2023-09-14 slides-herrsching2023.tex
typo

\begin{frame}{Proof translations}
  \textbf{Thm.} Every~\textsc{iqc}-proof remains correct, with at most a
  polynomial increase in length, if throughout we
  replace
  \[\begin{array}{rcl@{\quad\text{where}\quad}rcl}
    \exists & \leadsto & \exists_\mathrm{class},
    & \exists_\mathrm{class} &\defeqv& \neg\neg\exists, \\
    \vee & \leadsto & \vee_\mathrm{class},
    & \alpha \vee_\mathrm{class} \beta &\defeqv& \neg\neg(\alpha \vee \beta), \\
    = & \leadsto & =_\mathrm{class},
    & s =_\mathrm{class} t &\defeqv& \neg\neg(s = t).
  \end{array} \]
  \pause

@@

  \footnotesize
  \begin{tabular}{@{}lp{0.27\textwidth}p{0.48\textwidth}}
    \toprule
    forcing notion & statement about~$V^\nabla$ & external meaning \\
    \midrule
    surjection $\NN \twoheadrightarrow X$ &
    ``the gen.\@ surj.\@ is surjective'' &
    $\forall(a{\in}X)\_ \forall(\sigma{\in}X^*)\_ \nabla(\tau{\preccurlyeq}\sigma)\_ \exists(n{\in}\NN)\_ \tau[n] = a$. \\[0.5em]
    \pause

@@

    frame of opens &
    ``every complex number has a square root'' &
    For every open~$U \subseteq X$ and every cont.\@
    function $f : U \to \CC$, there is an open covering $U = \bigcup_i U_i$ such
    that for each index~$i$, there is a cont.\@ function $g : U_i \to \CC$
    such that~$g^2 = f$. \\[4.5em]

    Zariski &
    ``$x \neq 0 \Rightarrow \text{$x$ inv.}$'' &
    if the only f.p.\@ $k$-algebra in which~$x = 0$ is the zero algebra,
    then~$x$ is invertible in~$k$
  \end{tabular}
\end{frame}

@@

>>>>

\begin{frame}{Proof translations}
  \textbf{Thm.} Every~\textsc{iqc}-proof remains correct, with at most a
  polynomial increase in length, if throughout we
  replace
  \[\begin{array}{rcl@{\quad\text{where}\quad}rcl}
    \exists & \leadsto & \exists^\mathrm{cl},
    & \exists^\mathrm{cl} &\defeqv& \neg\neg\exists, \\
    \vee & \leadsto & \vee^\mathrm{cl},
    & \alpha \vee^\mathrm{cl} \beta &\defeqv& \neg\neg(\alpha \vee \beta), \\
    = & \leadsto & =^\mathrm{cl},
    & s =^\mathrm{cl} t &\defeqv& \neg\neg(s = t).
  \end{array} \]
  \pause

@@

  \footnotesize
  \begin{tabular}{@{}lp{0.27\textwidth}p{0.48\textwidth}@{}}
    \toprule
    forcing notion & statement about~$V^\nabla$ & external meaning \\
    \midrule
    surjection $\NN \twoheadrightarrow X$ &
    ``the gen.\@ surj.\@ is surjective'' &
    $\forall(a{\in}X)\_ \forall(\sigma{\in}X^*)\_ \nabla(\tau{\preccurlyeq}\sigma)\_ \exists(n{\in}\NN)\_ \tau[n] = a$. \\[0.5em]
    \pause

@@

    frame of opens &
    ``every complex number has a square root'' &
    For every open~$U \subseteq X$ and every cont.\@
    function $f : U \to \CC$, there is an open covering $U = \bigcup_i U_i$ such
    that for each index~$i$, there is a cont.\@ function $g : U_i \to \CC$
    such that~$g^2 = f$. \\[4.3em]

    big Zariski &
    ``$x \neq 0 \Rightarrow \text{$x$ inv.}$'' &
    If the only f.p.\@ $k$-algebra in which~$x = 0$ is the zero algebra,
    then~$x$ is invertible in~$k$.
  \end{tabular}
\end{frame}

@@

außerdem: Konservativität.
